---
output:
    html_document:
        code_folding: show
---
<br>

### Set environment
```{r message=0}
library(modeest)
library(tidyr)
library(kableExtra)
```
<br>

### Define 2 sets of data
```{r}
set1 <- c(10,2,3,2,4,2,5)
set2 <- c(20,12,13,12,14,12,15)
```
<br>

### Central Tendency: compute mean, medium, mode 
```{r}
ct_set1 <- c( mean(set1), median(set1), mfv(set1) )
ct_set2 <- c( mean(set2), median(set2), mfv(set2) )
```
<br>

### Variation: compute range, interquartile, variance, & standard deviation
```{r}

v_set1 <- round( c(range(set1), IQR(set1), var(set1), sd(set1) ), 2 )
v_set2 <- round( c(range(set2), IQR(set2), var(set2), sd(set2) ), 2 )
```
<br>

### Results: Central Tendency
```{r}
ct <- as.data.frame(rbind(ct_set1,ct_set2) )
colnames(ct) <- c("Mean","Median","Mode")
# print
ct %>% kbl(align = "ccc") %>%
  kable_styling(full_width = F)
```
<br>

### Results: Variation
```{r}
v1 <- as.data.frame( rbind(v_set1,v_set2) )
# combine the 2 columns of 'range' into 1
v2 <- v1 %>%
    unite("tmp", V1:V2, sep="-")
colnames(v2) <- c("Range","Interquartile","Variance", "Standard Deviation")
# print
v2 %>% kbl(align = "cccc") %>%
  kable_styling(full_width = F)
```
<br>

### Discussion
Central Tendency and Variation provide 2 ways to describe data.
The measures of center show the datasets' difference in their
numeric values. The measures of variation broadly show the degree to which they vary; it seems to hints at little variation - only the range
provides a clue what the underlying values may look like.  
<br>
The coefficient of variation can be used to compare the datasets:
$CV = (\frac{S}{\bar x} ) \cdot 100$  
\ $CV_{set1} =$ `r sprintf("%1.2f%%", (sd(set1)/mean(set1)) * 100 ) `  
\ $CV_{set2} =$ `r sprintf("%1.2f%%", (sd(set2)/mean(set2)) * 100 ) `  
<br>
While both datasets have the same standard deviation, they differ relative
to their mean.  
<br><br>

### GitHub
Related file(s) can be found at <a href="https://github.com/eltownes/USF-LIS4273/tree/master/Module%202">Git Me</a>
<br><br>