
---
output:
    html_document:
        code_folding: hide 
---
<br>

### Set environment
```{r message=0, results='hide'}
env <- c("dplyr","readxl","tidyr","corrgram")
lapply(env, library, character.only = 1)
```
<br>

### <span style="color:rgb(219,228,66)">Part I</span>
<br>
**The director of manufacturing at a cookie factory needs to determine whether 
a new machine is producing a particular type of cookie according to the 
manufacturer's specifications, which indicate that cookies should have a 
mean breaking strength of 70 and standard deviation of 3.5 pounds. 
A sample of 49 of cookies reveals a mean breaking strength of 69.1 pounds.**  

**1 - Hypothesis**

null hypothesis: the sample of 69.1 lbs is within the manufacturer's
specifications.  
alternative hypothesis:  the sample of 69.1 lbs is NOT within the 
manufacturer's specifications.  

This will be a two-tailed test because any sample mean can be either more or 
less than the population mean of 70 lbs with a standard deviation of 3.5
pounds - testing whether the cookies are either too soft or too hard.

$H_{0}$ : $\mu$ = 69.1  
$H_{a}$ : $\mu \ne$ 69.1  
<br>
"If z is in the upper tail (z > 0), compute the probability that z is greater than or equal to the value of the test statistic. If z is in the lower tail
(z < 0), compute the probability that z is less than or
equal to the value of the test statistic. Double the tail area to obtain
the p–value."  
<br>

**2 - Within acceptable limits?**

Is there evidence that the machine is not meeting the manufacturer's specifications for average strength? Use a 0.05 level of significance.  

$\bar{x} = 69.1 \ ->$ `xbar` `r xbar <- 69.1`  
$\mu = 70 \ ->$ `mu` `r mu <- 70`  
$\sigma = 3.5 \ ->$ `sd` `r sd <- 3.5`  
$n = 49 \ ->$ `n` `r n <- 49` 
$\alpha = 0.05$ `alpha` `r alpha <- 0.05`

$\Large z = \frac {\bar{x}-\mu}{\sigma / \sqrt {n}} =$
$\Large \frac{`r xbar` - `r mu`}{`r sd` / \sqrt `r n`} =$
`r z <- (xbar-mu) / (sd/sqrt(n) )` $`r z`$  
<br>

With z = `r z`, the two-tailed p-value is `r pval <- 2 * pnorm(z)` `r pval`  

Rejection?  
p-value $\leq$ alpha ? ---> `r pval` $\leq$ `r alpha` = **`r pval <= alpha`**  
<br>
**Do NOT reject $H_{0}$**, the sample of `r xbar` lbs is within the manufacturer's
specifications.  
<br>

**3 - What if...**

the standard deviation were specified as 1.75 pounds.  

$\sigma = 1.75 \ ->$ `sd` `r sd <- 1.75`  

$\Large z = \frac {\bar{x}-\mu}{\sigma / \sqrt {n}} =$
$\Large \frac{`r xbar` - `r mu`}{`r sd` / \sqrt `r n`} =$
`r z <- (xbar-mu) / (sd/sqrt(n) )` $`r z`$  
<br>

With z = `r z`, the two-tailed p-value is `r pval <- 2 * pnorm(z)` `r pval`  

Rejection?  
p-value $\leq$ alpha ? ---> `r pval` $\leq$ `r alpha` = **`r pval <= alpha`**  
<br>
**Do reject $H_{0}$**, the sample of `r xbar` lbs is not within the manufacturer's
specifications.  
<br>

**4 - What if...**

the sample mean were 69 pounds and the standard deviation is 3.5 lbs.  

$\bar{x} = 69 \ ->$ `xbar` `r xbar <- 69`  
$\sigma = 3.5 \ ->$ `sd` `r sd <- 3.5`  

$\Large z = \frac {\bar{x}-\mu}{\sigma / \sqrt {n}} =$
$\Large \frac{`r xbar` - `r mu`}{`r sd` / \sqrt `r n`} =$
`r z <- (xbar-mu) / (sd/sqrt(n) )` $`r z`$  
<br>

With z = `r z`, the two-tailed p-value is `r pval <- 2 * pnorm(z)` `r pval`  

Rejection?  
p-value $\leq$ alpha ? ---> `r pval` $\leq$ `r alpha` = **`r pval <= alpha`**  
<br>
**Do reject $H_{0}$**, the sample of `r xbar` lbs is NOT within the manufacturer's
specifications.  
<br>

### <span style="color:rgb(219,228,66)">Part II</span>
<br>
**If x̅ = 85, σ = 8, and n = 64, set up 95% confidence interval estimate 
of the population mean μ.**  

For CI using: 
$\Large \bar{x} \ \pm z \ \frac{alpha}{2} \cdot \frac{\sigma}{sqrt(n)}$  
<br>

The 95% Confidence Interval:
<br>
<br>
$\bar{x} = 85 \ ->$ `xbar` `r xbar <- 85`  
$alpha = 0.05 \ ->$ `alpha` `r alpha <- 0.05`  
$\sigma = 8 \ ->$ `sd` `r sd <- 8`  
$n = 64 \ ->$ `n` `r n <- 64`  
<br>
In R: 
`xbar +- qnorm( alpha/2 ) * ( sd/sqrt(n) ) `  
`r v1 <- xbar + qnorm( alpha/2 ) * ( sd/sqrt(n) )` Value 1 = `r v1`  
`r v2 <- xbar - qnorm( alpha/2 ) * ( sd/sqrt(n) )` Value 2 = `r v2`  
<br>
The 95% CI is ( `r round(v1-0.5,0)` , `r round(v2+0.5,0)` ).  
<br>


<br>

### <span style="color:rgb(219,228,66)">Part III</span>
<br>

The girls and boys data sets:
```{r message=FALSE}
girls <- ( read_excel("Mod 8 dataset.xls", n_max = 5 ) )
boys <- ( read_excel("Mod 8 dataset.xls", skip = 7 ) )

# tidy the data
colnames(girls) <- c("category","n1","n2","n3")
girls$id <- "girls"
girls <- girls %>%
    select(id,everything() )
colnames(boys) <- c("category","n1","n2","n3")
boys$id <- "boys"
boys <- boys %>%
    select(id,everything() )
```

```{r}
girls
boys
```


```{r }
# prep for cor test
gd <- girls %>%
    slice_tail(n=1) %>%
    select(n1,n2,n3) 
bd <- boys %>%
    slice_tail(n=1) %>%
    select(n1,n2,n3)
c <- cor( as.numeric(gd) , as.numeric(bd) )

```
The correlation coefficient on the totals is `r c`  
<br>

The highly correlated correlogram:
```{r }
# prep for corrgram
all <- rbind( slice_head(girls,n=4) , slice_head(boys,n=4) )
cg <- corrgram( all, order=TRUE, lower.panel=panel.ellipse,
  upper.panel=panel.pts, text.panel=panel.txt )
```
<br><br>


### GitHub
Related file(s) can be found at <a href="https://github.com/eltownes/USF-LIS4273/tree/master/Module%208">Git Me</a>
<br><br>

