
---
output:
    html_document:
        code_folding: hide 
---
<br>

### Set environment
```{r message=0, results='hide'}
env <- c("tidyverse")
lapply(env, library, character.only = 1)
```
<br>

**Report on drug and stress level by using R. Provide a full summary report 
on the result of ANOVA testing and what does it mean.**  
<br>

The Stress dataset.  
<br>
A researcher is interested in the effects of drug against stress reaction. 
She gives a reaction time test to three different groups of subjects: one 
group that is under a great deal of stress, one group under a moderate 
amount of stress, and a third group that is under almost no stress. The 
subjects of the study were instructed to take the drug test during their 
next stress episode and to report their stress on a scale of 1 to 10 
(10 being most pain).

```{r }
testOrig <- tibble(
    H = c(10,9,8,9,10,8),
    M = c(8,10,6,7,8,8), 
    L = c(4,6,6,4,2,2) 
)
testOrig
```
<br>

Exploratory Data Analysis
```{r}
boxplot(testOrig)
summary(testOrig)
```
<br>

Pivot data for analysis
```{r }
testR <- pivot_longer( testOrig , cols = everything() ,
                       names_to = "type" ,
                       values_to = "level" )
testR$type <- as.factor( testR$type )
head( testR )
```
<br>

The ANOVA test:
```{r }
aovOut <- aov( level ~ type , data = testR )
summary( aovOut  )
```
The test shows:  
F(2,15) = 21.36 and P < 0

As the F ratio is not near 1.0 and the P-value is negative, it shows an 
imbalance and suggests that a differences exists between 
the H, M, and L Stress types. A null hypothesis can be rejected. The samples
indicate that for the general population there are mean differences among the treatments being compared.  

However, the ANOVA simply establishes that differences exist, it does not 
indicate exactly which treatments are different.

A Tukey Honestly Significant Difference (HSD) can compare the treatments,
two at a time, to test the significance of the mean differences.

The Tukey HSD test:
```{r }
TukeyHSD( aovOut )

```
The test shows that the greatest difference lies with the L-H pair, 
followed by the M-L pair. The M-H pair shows no significant difference. This
support an intuitive interpretation of the above boxplot.
<br><br>


### GitHub
Related file(s) can be found at <a href="https://github.com/eltownes/USF-LIS4273/tree/master/Module%210">Git Me</a>
<br><br>

