
---
output:
    html_document:
        code_folding: hide 
---
<br>

### Set environment
```{r message=0, results='hide'}
env <- c("dplyr")
lapply(env, library, character.only = 1)
```
<br>

### <span style="color:rgb(219,228,66)">Contingency Table</span>
<br>

#### Create table
```{r }
data1 <- matrix( c(10,20,20,40), nrow=2, byrow=1,
                 dimnames=list( c("A","A1"), c("B","B1") ) )
table1 <- as.table( data1 )
table1
```
<br>

#### Calculate relative frequencies
```{r }
propTable <- prop.table(table1)
propTable
```
<br>

#### Find probabilities for 'A' and 'B'

* P(A) : sum the table row  
`r pA <- propTable %>% rowSums() %>% .["A"] ` <!--store-->
` pA <- propTable %>% rowSums() %>% .["A"] ` = `r pA`
<br>
    
* P(B) : sum the table column  
`r pB <- propTable %>% colSums() %>% .["B"] ` <!--store-->
` pB <- propTable %>% colSums() %>% .["B"] ` = `r pB`
<br>

* P(A and B) : is already on the rel freq table as $P(A) \cap P(B)$ but we can
still calculate  
` pA * pB ` = `r pA * pB `
<br>

* P(A or B) : P(A) + P(B) - P(A and B)  
` pA + pB - (pA * pB) ` = `r pA + pB - (pA * pB) `
<br>

* P(A or B) disjoint : P(A) + P(B)  
` pA + pB ` = `r pA + pB `  
<br><br>


### <span style="color:rgb(219,228,66)">Bayes Theorem</span>

Jane is getting married tomorrow, at an outdoor ceremony in the desert. In recent years, it has rained only 5 days each year. Unfortunately,
the weatherman has predicted rain for tomorrow. When it actually rains, the weatherman correctly forecasts rain 90% of the time. When it
doesn't rain, he incorrectly forecasts rain 10% of the time.  

**What is the probability that it will rain on the day of Jane's wedding?**  
<br>

#### Discussion

Calculate the probabilities we know. 
( R - rain, NR - no rain, F - weatherman forecast )  

* Relative  
    * P(R) : ` pR <- 5/365 ` = `r pR <- 5/365 ` `r pR `  
    * P(NR) : ` pNR <- 360/365 ` = `r pNR <- 360/365 ` `r pNR `  
* Conditional  
    * P(F|R) : 0.9 ` pFR <- 0.9 ` `r pFR <- 0.9 `  
    * p(F|NR) : 0.1 ` pFNR <- 0.1 ` `r pFNR <- 0.1 `  
<br>

Proportional probability would suggest that there's a `r round(pR,2)` or 
`r sprintf("%1.2f%%", pR*100) ` chance that it will rain on Jane's wedding day.  
<br>
Given the weatherman's prediction history, we can use Bayes to
update our initial probability - the prior. We now want to predict the 
probability of rain based on the new forecast information - the 
posterior : $P(R|F)$  

<br>
Using Bayes :
$\large P(R|F) = \frac{P(F|R) \cdot P(R)}{\sum P(F)}$ = 
$\large \frac{P(F|R) \cdot P(R)} {P(F|R) \cdot P(R) + P(F|NR) \cdot P(NR)}$  
<br>

`r b <- pR*pFR/(pR*pFR+pNR*pFNR) `
Substituting results in :
$\large \frac{`r pFR ` \cdot `r pR `}
    {`r pFR ` \cdot `r pR ` + `r pFNR ` \cdot `r pNR `}$ = `r b`  
<br>

Interpretation:  

* The weatherman's forecast increased our probability of rain from `r sprintf("%1.2f%%", pR*100) ` to `r sprintf("%1.2f%%", b*100) `. A 
big increase but overall, there's a low probability of rain on 
Jane's wedding day.  
* While the weatherman's rain prediction rate is high, the fact that it only
rains a few days per year, significantly lowers his **daily** probability
accuracy. Even if the weatherman had an accurate rain forecast rate of 99%,
the probability of rain on the wedding day would still be only about
0.58 or 58% - slightly better than a coin toss.
<br><br>

### GitHub
Related file(s) can be found at <a href="https://github.com/eltownes/USF-LIS4273/tree/master/Module%204">Git Me</a>
<br><br>

