
---
output:
    html_document:
        code_folding: hide 
---
<br>

### Set environment
```{r message=0, results='hide'}
env <- c("dplyr", "ggplot2", "glue")
lapply(env, library, character.only = 1)
```
<br>

### Load assigned dataset
```{r message=0}
boardingScreeners <- c(287,243,237,227,247,264,247,247,251,254,
                       277,303,285,254,280,264,261,292,248,253)
securityViolations <- c(271,261,230,225,236,252,243,247,238,274,
                        256,305,273,234,261,265,241,292,228,252)
secData <- tibble( boardingScreeners,securityViolations )
```
<br>

### Create EDA scatter plot
```{r}
plot(secData)
```
<br>
The plot shows a clear positive relationship: security violations (y) increase as
more screening agents (x) are used. One would expect any calculated 
correlation value to be positive and nearer to one (1) than zero (0).
<br>

### Calculate Pearson’s Sample correlation coefficient
```{r}
corPearson <- cor.test(secData$boardingScreeners, secData$securityViolations,
                       method="pearson")
corPearson
```
<br>

### Calculate Spearman’s Rank Coefficient
```{r}
corSpearman <- cor.test(secData$boardingScreeners, secData$securityViolations,
                        method="spearman")
corSpearman
```
<br>

### Describe the association between boarding screeners and security violations
Correlation measures both the strength of association between two
variables and the direction of that relationship. The scale ranges from
-1 to +1 and the result is expressed as a decimal value. A value of 0 typically
means that there is no relationship.  
<br>
The Pearson correlation is the most cited measure when continuous and linearly related variables are involved. For the above dataset, the r value is 0.84 which 
shows a strong positive correlation between screeners and violations.  
<br>
The Spearman correlation is a non-parametric test. In essence it shows a rank
correlation which can be between both continuous or ordinal variables. In
the case of continuous data, the raw values are ranked - they become 
rank-ordered variables. For the dataset, the Spearman correlation comes
in at 0.76. Like the Pearson score, it shows a strong and positive association.  
<br>
While a graph for a Pearson correlation will show a constant rate of change, 
the Spearman describes a monotonic relationship where the rate of change
can vary. Lastly, as the ref points out, when the linearity of the data is in question it may be best to opt for the Spearman correlation.  
<br>

Ref:  
[Towards Data Science - Clearly explained: Pearson vs Spearman Correlation Coefficient](https://towardsdatascience.com/clearly-explained-pearson-v-s-spearman-correlation-coefficient-ada2f473b8#:~:text=The%20fundamental%20difference%20between%20the,with%20monotonic%20relationships%20as%20well.)
<br>

### Prep report
```{r message=0 }
# prep label
corLabel <- glue("
    Correlation Coeficients
    Pearson: {round(corPearson$estimate,2)}
    Spearman: {round(corSpearman$estimate,2)}
    ")
ggplot(secData, aes(boardingScreeners,securityViolations) ) +
    geom_point() +
    geom_smooth(method="lm", se=0, aes(colour="lm")) +
    geom_smooth(method="loess", se=0, aes(colour="loess")) +
    geom_label(label = corLabel, aes(240,295) ) +
    labs(title = "Airport Security Screening", subtitle = "(1988-1999)",
         x = "Screeners", y = "Violations  ", colour = "Model") +
    theme_classic() + 
    theme( plot.title = element_text(hjust=0.5),
           plot.subtitle = element_text(hjust=0.5),
           axis.title.x = element_text(vjust=-1),
           axis.title.y = element_text(angle = 0, vjust=0.5)
           )
```


<br><br>


### GitHub
Related file(s) can be found at <a href="https://github.com/eltownes/USF-LIS4273/tree/master/Module%203">Git Me</a>
<br><br>

