
---
output:
    html_document:
        code_folding: hide 
---
<br>

### Set environment
```{r message=0, results='hide'}
env <- c("dplyr","tibble")
lapply(env, library, character.only = 1)
```
<br>

<br>
The dataset is from a study whose research hypothesis was:  

* "There will be a difference between boys and girls in the number of times
they raise their hand in class."  

For the purposes of this study, boys were coded as 2 on gender and
girls were coded as 1.  

The dataset:  
```{r }
classroom <- tribble(
    ~gender, ~handup,
    1,9,
    2,11,
    2,1,
    1,8,
    2,2,
    2,6,
    1,3,
    2,4,
    1,8,
    2,3,
    1,10,
    2,6
)
classroom
```
<br>

Compute t-test.  
```{r }
boys <- classroom %>%
    filter( gender == 1 ) %>%
    select( -gender )
girls <- classroom %>%
    filter( gender == 2 ) %>%
    select( -gender )
res <- t.test( boys , girls )
res
```
<br>

1] The means:  
x, boys: `r res[["estimate"]][["mean of x"]] `  
y, girls: `r res[["estimate"]][["mean of y"]] `  
<br>

2] Degree of freedom:  
`r res[["parameter"]][["df"]] `  
<br>

3] T-test statistic:  
`r res[["statistic"]][["t"]] `  
<br>

4] P-value:  
`r pVal <- res[["p.value"]] ` `r pVal `  
<br>

5] At alpha value of 0.05  
Is p-value $\le$ 0.05? `r pVal <= 0.05 `  
Do **NOT** reject the null hypothesis. There is a difference.  
<br>

6] What critical value would your obtained t-test value have to exceed
to be significant at the 0.01 level? - assume a two-tailed test.  

Formula: `qt( 0.01/2 , 'degrees of freedom' , lower.tail=FALSE) `  

```{r }
df <- res$parameter
t_crit <- qt( 0.01/2 , df , lower.tail = FALSE )
```

The critical value is `r t_crit `.  
Given a two-tailed test, the critical values would have 
to be: $\lt$ `r t_crit*-1 ` and $\gt$ `r t_crit `.


<br><br>


### GitHub
Related file(s) can be found at <a href="https://github.com/eltownes/USF-LIS4273/tree/master/Module%209">Git Me</a>
<br><br>

