
---
output:
    html_document:
        code_folding: hide 
---
<br>

### Set environment
```{r message=0, results='hide'}
env <- c("dplyr")
lapply(env, library, character.only = 1)
```
<br>

### <span style="color:rgb(219,228,66)">Part A</span>
<br>


#### Create dataset
```{r }
data1 <- c(8,14,16,10,11) ; data1
```
<br>

(a) ##### Compute population mean & standard deviation
```{r }
pop_mu <- mean(data1)
pop_sd <- sd(data1)
```

Mean is `r pop_mu ` and the standard deviation is `r pop_sd `  
<br>

```{r }
fun_calc <- function(trials){

    
# ---------- get samples
    
# sampling and replicate
s1_rs <- replicate( trials, sample(data1,2,replace=1) )
# mean & sd
s1_xbar <- apply(s1_rs,2,mean)
s1_sd <- apply(s1_rs,2,sd)
# function for standard error of the mean
fun_sem <- function(x) sd(x)/sqrt(length(x))
s1_sem<- apply(s1_rs, 2, fun_sem)
# format table
s1_tbl <- as.data.frame( rbind(s1_rs, s1_xbar, s1_sd, s1_sem) )
colnames(s1_tbl) <- c( paste("Trial ", seq(trials) ) )
rownames(s1_tbl) <- c("observation 1","observation 2","xbar", "std dev",
                      "std error")
# view filter
if(trials <= 4){
    viewTbl <- s1_tbl
} else {
    test <- data.frame("-----")
    colnames(test)<-   "-----"
    viewTbl <- bind_cols(s1_tbl[, 1:2], test, s1_tbl[, (trials-1):trials] )
}
# view table
print(viewTbl)


# ---------- roll up

# calculate sampling values - population std error mean is NA
pop_sem <- NA
rs_xbar <- mean(s1_rs)
rs_sd <- sd(s1_rs)
rs_sem <- sd(s1_rs) / sqrt(length(s1_rs))
# assemble table
pop_rollUp <- as.data.frame( rbind(pop_mu, pop_sd, pop_sem) )
rs_rollUp <- as.data.frame( rbind(rs_xbar, rs_sd, rs_sem) )
rollup <- bind_cols(pop_rollUp, rs_rollUp  )
colnames(rollup) <- c("Population","Samples")
rownames(rollup) <- c("mean/xbar","std dev","std err")
# view table
print(rollup)
}
```

`r trials <- 1`  
(b) ##### Random sample(s) of 2 from the population - **`r trials` trial(s)**
(c) ##### Compute the mean & standard deviations
(d) ##### Comparing the population and the samples
```{r message=0}
fun_calc(trials)
```
<br>


`r trials <- 100`  
(b) ##### Random sample(s) of 2 from the population - **`r trials` trial(s)**
(c) ##### Compute the mean & standard deviations
(d) ##### Comparing the population and the samples
```{r message=0}
fun_calc(trials)
```
<br>


### <span style="color:rgb(219,228,66)">Part B</span>
<br>


**Suppose that the sample size n = 100 and the population proportion p = 0.95.**
`r n <- 100; p <- 0.95 `  
<br>

(a) Does the sample proportion p have approximately a normal distribution? Explain.

**No.**

One rule states: "the sample size is considered large enough if the expected number of successes and failures are both at least 10: 
$np \ge 10$ and $n(1-p) \ge 10$  .

Given the data : $np \ge 10$ = `r n*p >= 10 ` 
    and $n(1-p) \ge 10$ = `r n*(1-p) >= 10 `  
<br>


(b) What is the smallest value of n for which the sampling distribution of p is approximately normal?  

**201.**

Solving each sub-rule for n means
n $\ge$ `r 10/p ` and n $\ge$ `r 10/(1-p)` .


<br><br>


### GitHub
Related file(s) can be found at <a href="https://github.com/eltownes/USF-LIS4273/tree/master/Module%205">Git Me</a>
<br><br>

