
---
output:
    html_document:
        code_folding: hide 
---
<br>

### Set environment
```{r message=0, results='hide'}
env <- c("tidyverse")
lapply(env, library, character.only = 1)
```
<br>

**The Hotels data table**
```{r }
hotels <- tribble(
    ~Same_Hotel_Again,~Beachcomber, ~Windsurfer,
    "Yes",163,154,
    "No",64,108
)
hotels
```
<br>

**Chi-Square test for the difference between two proportions:**  
<br>
$H_{0}$ : $\pi_1$ = $\pi_2$ (The satisfaction of Beachcomber guests is 
equal to those of Windsurfer guests)  
$H_{a}$ : $\pi_1$ $\ne$ $\pi_2$ (The satisfaction of Beachcomber guests is NOT
equal to those of Windsurfer guests)  
<br>

**Prep data and calculate.**  
```{r }
tabHotel <- hotels[,2:3]
tabHotel

chiHotel <- chisq.test( tabHotel )
chiHotel
```
<br>

**Assuming a significance of 0.05**  
<br>

The $x^2_{0.05}$ with 1 degree of freedom equals ...
```{r }
zAlpha <- qchisq( c(0.05), df=1, lower.tail=FALSE)
zAlpha
```
<br>

**Analyze the data**
```{r }
ggplot( data.frame(x = 0:10) , aes(x = x) ) +
    stat_function(fun = df,
                  args = list( df1=1 , df2=Inf ),
                  col = "red"
                  ) +
    geom_vline( xintercept=zAlpha ) +
        geom_label( aes(label="alpha", x=zAlpha, y=1.5 ) )  +
    geom_vline( xintercept=chiHotel$statistic ) +
        geom_label( aes( label="stat", x=chiHotel$statistic, y=1.5 ) ) + 
    labs(title = "Chi-Square Test",
         x = "x", y = "y") +
    theme( plot.title = element_text(hjust=0.5),
           axis.title.x = element_text(vjust=-1),
           axis.title.y = element_text(angle = 0, vjust=0.5)
        )

```

**Decision Rule:**  
If $x^2_{STAT} \gt x^2_{\alpha}$ reject $H_{0}$ otherwise,
do not reject $H_{0}$  
<br>

We can reject the null hypothesis.  
At a significance level of 0.05, the satisfaction of Beachcomber guests is NOT equal to those 
of Windsurfer guests.  

```{r }
pvtHotel <- pivot_longer( hotels , cols=c(Beachcomber,Windsurfer) )

ggplot( pvtHotel, aes(x = Same_Hotel_Again, fill = value ) ) +
    geom_bar( aes(weight = value) ) +
    facet_wrap( ~name )
```
<br>  

By examining the bar graph of the provided data and at a significance level of 0.05, we can conclude that the Beachcomber is definitely the preferred hotel.
<br><br>


### GitHub
Related file(s) can be found at <a href="https://github.com/eltownes/USF-LIS4273/tree/master/Module%2011">Git Me</a>
<br><br>

