
---
output:
    html_document:
        code_folding: hide 
---
<br>

### Set environment
```{r message=0, results='hide'}
env <- c("tidyverse","forecast")
lapply(env, library, character.only = 1)
```
<br>

**The credit card charges data**  
<br>
A quick EDA of the data.
```{r }
visa <- c(31.9,27,31.3,31,39.4,40.7,42.3,49.5,45,50,50.9,58.5,39.4,
            36.2,40.5,44.6,46.8,44.7,52.2,54,48.8 ,55.8,58.7,63.4)
visaTS <- ts(visa, frequency=12, start=c(2018,1) )
visaTS
summary( visaTS )
autoplot( visaTS )
```
<br>

**Apply smoothing**  
<br>

We will use simple exponential smoothing by using the HoltWinters function.
The EDA hinted at an intuitive seasonal pattern where more charges occur during
the typical holiday seasons so, we'll "turn on" the gamma argument.  

```{r message=FALSE}
visaXMA <- HoltWinters(visaTS, beta=FALSE, gamma=TRUE)
visaXMA

autoplot( visaXMA$x, size=1 ) +
    autolayer( visaXMA$fitted, color="Red" ) +
    labs(title = "Visa Charges",
         subtitle = "(With HoltWinter Seasonal Smoothing)",
         x = "Year", y = "Charges $" ) +
    theme_classic() + 
    theme(legend.position = "none") +
    theme( plot.title = element_text(hjust=0.5),
           plot.subtitle = element_text(hjust=0.5, color="Red"),
           axis.title.x = element_text(vjust=-1),
           axis.title.y = element_text(angle = 0, vjust=0.5)
           )

```
<br>

**Forecasting**  
<br>

Let's forecast the next 12 months using a confidence level of 50%, 75%, 
and 90%.  

```{r}
visaFC <- forecast(visaXMA, h=12, level=c(50,75,90) )
visaFC

autoplot( visaFC ) +
    labs(title = "Charge Prediction for 2020",
         subtitle = "(50%, 75%, 90%)", x = "Year", y = "Charges $" ) +
    theme_classic() + 
    theme(legend.position = "none") +
    theme( plot.title = element_text(hjust=0.5),
           plot.subtitle = element_text(hjust=0.5, color="Blue"),
           axis.title.x = element_text(vjust=-1),
           axis.title.y = element_text(angle = 0, vjust=0.5)
           )

```



<br><br>


### GitHub
Related file(s) can be found at <a href="https://github.com/eltownes/USF-LIS4273/tree/master/Module%2012">Git Me</a>
<br><br>

